Parsing Postfix log files

### 2007-10-08 doc/logparser.tex
Minor graph changes

% Useful stuff for math mode.
\usepackage{amstext}
% Wrapping of URLs.  This is used in the bibliography instead of \url{}.
\usepackage{url}
%\DeclareUrlCommand\urlwrap{}
%\newcommand{\urlwrapfill}[1]{\urlwrap{#1}\hfill}
\newcommand{\parsername}{ASO}
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib]{tocbibind}
% When creating a PDF make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author don't work - I think either gnuplot or graphviz clobbers it.
\usepackage{hyperref}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0 0 0},
%    breaklinks  = true,
}
\usepackage{breakurl}

@@

Parser coverage is divided into two topics in this section: log lines
covered, and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and not intrusive:
adding new rules or modifying existing rules is simplified by the parser
implementation.  Improving the logical coverage is harder, as the actions
taken by Postfix must be reconstructed, and the new sequence of actions
integrated into the existing model without breaking the existing sequences.
detecting a deficiency in the parsing algorithm is also significantly
harder than detecting unparsed log lines, as the parser will warn about any
unparsed line, whereas the author must spot a flaw in the parser by
understanding the warnings generated.

@@

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number lines}.  The explanation
for this took some time to arrive at, but it turns out to be reasonably
simple.  Examining the database after parsing logfile~22 in isolation shows
64 rules matched during parsing, but as shown in graph~\ref{log22
distribution by rule} the distribution of matching rules is extremely
skewed.  Comparing the number of log lines per program versus rules per
program in graph~\ref{log22 line and rule distribution by program} shows
that \texttt{postfix/qmgr} has the highest number of log lines, however it
has a low number of rules.  This combination results in a low average
number of rules to try for \texttt{postfix/qmgr} log lines, reducing the
overall average number of rules to try per log line.  In turn this reduces
the average parsing time per log line, and thus the parsing time for the
log file.  While the parsing time is substantially greater than the time
for the log files on either side, the increase from optimal ordering to
reversed ordering is proportionally smaller, producing the dips in
graph~\ref{normal vs shuffled vs reversed ordering factor}.

@@

\subsubsection{Log 22 line and rule distribution by program}
\label{log22 line and rule distribution by program}

\noindent
\includegraphics{build/plot-log22-distribution-versus-rule-count.ps}

\subsubsection{Log 22 lines per rule}
\label{log22 lines per rule}

As shown in this graph, the most common log lines (see graph~\ref{log22
line and rule distribution by program}) are from programs which have a low
number of rules, thus reducing the average parsing time per line, and the
overall execution time relative to the input size.

\noindent
\includegraphics{build/plot-log22-distribution-versus-rule-count-factor.ps}

>>>>

% Useful stuff for math mode.
\usepackage{amstext}
% Wrapping of URLs.  This is used in the bibliography instead of \url{}.
\usepackage{url}
%\DeclareUrlCommand\urlwrap{}
%\newcommand{\urlwrapfill}[1]{\urlwrap{#1}\hfill}
\newcommand{\parsername}{ASO}
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib]{tocbibind}
% When creating a PDF make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author don't work - I think either gnuplot or graphviz clobbers it.
\usepackage{hyperref}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0 0 0},
%    breaklinks  = true,
}
\usepackage{breakurl}
\usepackage{lastpage}

@@

Parser coverage is divided into two topics in this section: log lines
covered, and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and not intrusive:
adding new rules or modifying existing rules is simplified by the parser
implementation.  Improving the logical coverage is harder, as the actions
taken by Postfix must be reconstructed, and the new sequence of actions
integrated into the existing model without breaking the existing sequences.
Detecting a deficiency in the parsing algorithm is also significantly
harder than detecting unparsed log lines, as the parser will warn about any
unparsed line, whereas the author must spot a flaw in the parser by
understanding the warnings generated.

@@

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number lines}.  The explanation
for this took some time to arrive at, but it turns out to be reasonably
simple.  Examining the database after parsing logfile~22 in isolation shows
64 rules matched during parsing, but as shown in graph~\ref{log22
distribution by rule} the distribution of matching rules is extremely
skewed.  Comparing the number of log lines per program versus rules per
program in graphs~\ref{log22 log line and rule distribution by program}
shows that \texttt{postfix/qmgr} has the highest number of log lines,
however it has a low number of rules.  This combination results in a low
average number of rules to try for \texttt{postfix/qmgr} log lines,
reducing the overall average number of rules to try per log line.  In turn
this reduces the average parsing time per log line, and thus the parsing
time for the log file.  While the parsing time is substantially greater
than the time for the log files on either side, the increase from optimal
ordering to reversed ordering is proportionally smaller, producing the dips
in graph~\ref{normal vs shuffled vs reversed ordering factor}.

@@

\subsubsection{Log 22 log line and rule distribution by program}
\label{log22 log line and rule distribution by program}

\noindent
\includegraphics{build/plot-log22-rules.ps}

\noindent
\includegraphics{build/plot-log22-log-lines.ps}

As shown in graphs above, the most common log lines are from programs which
have a low number of rules, thus reducing the average parsing time per
line, and the overall execution time relative to the input size.
